Stack Safety as a Security Property

### 2020-07-09 main.tex
done with pass through 5.3; releasing lock

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

@@

Both the integrity and confidentiality components of local state
encapsulation concern themselves with the accessibility of data, and
may hold on a wide range of programs that violate expected control
flow. Say, if {\tt g} managed to jump in the middle of the
code of {\tt h} with its call, confidentiality and integrity would
still be satisfied.
%
However, that is still unwanted behavior: for example, a caller (or a
callee) should not be able to bypass a password check by jumping (or
returning) in the middle of a block of code --- even if such an execution
would ostensibly satisfy local state encapsulation.
%
Stack safety is traditionally associated with preventing such unwanted
flows: every jump between functions should either be a call or a
return, and each return should be to the instruction after its
corresponding call.

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (call maps, function maps, return maps, and entry
maps): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

\definition{Control Separation}

A system enjoys {\em control separation} with regard to a function map
\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
belongs to different functions according to the function map,
either $\mach_1$ is a valid call or a valid return:

@@

For example, consider the unsafe substitution:
\[
5 : ~ \jalr ~ r_4 ~ \rra ~ 0
\]
This violates control separation as the instruction that causes a
function switch is not a valid call (or a return) according to the
call map.

\definition{Entry Integrity}

A system enjoys {\em entry integrity} with regard to a call map
\(\callmap\) and entry map \(\entmap\) if for any initial state
\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

The same unsafe substitution violates entry integrity
\[
5 : ~ \jalr ~ r_4 ~ \rra ~ 0
\]
as its target (instruction 42) is not an established entry point for
{\tt h} according to the entry map.

@@

\definition{Return Integrity}

\newcommand*{\Unmatched}{\mathit{Unmatched}}

Return integrity captures the expected behavior of marked returns with
regard to their call points. Given a call map $\callmap$ and a return
map $\retmap$, we begin by defining a coinductive relation
$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.

Formally:
\begin{itemize}
\item A trace that starts with a return is related to itself:

  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}

@@

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))
              \Unmatched ~ \MPT ~ \MPT'\)}
           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}

\item Finally, if a trace starts with the call, then we must discard the first
  unmatched return to get to the real one:

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state pair
\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
is the first unmatched return such that for some $\pol_r$
\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
then $\mach_r$ corresponds to a return state for $\mach_c$:
\[\ret{\mach_c}\ \mach_r\]

For example, if we added an instruction between 45 and 46:
\[
45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
\]
This would cause the return from {\tt h} to go to instruction 8 and
return to {\tt main} instead of {\tt g}. While the instruction that
causes the return would still be a valid return (satisfying control
separation), the state after the return would not correspond to an
$\Unmatched$ return for {\tt g}'s call to {\tt h}, violating return
integrity.

@@

The above properties are somewhat orthogonal, but together they
capture a standard notion of control flow for a stack: functions keep
their code separate outside of calls and returns, calls respect
explicit entry points, and returns respect implicit entry points
corresponding to their call.  A system that enjoys them with regard to
a code, a call, a return, and an entry map is said to enjoy
{\em well-bracketed control flow} with regard to those maps.

@@

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
  and stack confidentiality test $\testConf{\mach}{\mach'}$
  with regard to the original machine state. We write
  this $\testVS{(\mach, \pol)}{\vs}$.

@@

\definition
%
  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
  satisfies the eager testing property at depth 1 with regard to the initial
  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
%
  \rb{I don't think the initial variant has been defined, but it's
    straightforward.}

@@

      A system enjoys {\em observable stack integrity} with regard to
      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with regard to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with regard to a call map
      and a function map if it enjoys both observable stack integrity and
      observable stack confidentiality with regard to them.

>>>>

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\mpstatename \in \MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with respect to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

@@

Both the integrity and confidentiality components of local state
encapsulation concern themselves only with the accessibility of data, and
may still hold on a wide range of programs that violate expected control
flow. Say, if {\tt g} managed to jump in the middle of the
code of {\tt h} with its call, confidentiality and integrity would
still be satisfied.
%
However, that is still unwanted behavior: for example, a caller (or a
callee) should not be able to bypass a password check by jumping (or
returning) in the middle of a block of code --- even if such an execution
would ostensibly satisfy local state encapsulation.
%
Informal definitions of stack safety usually \apt{evidence?} aim to prevent such unwanted
flows: every jump between functions should either be a call or a
return, and each return should be to the instruction after its
corresponding call.

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (call map $\callmap$, function map $\codemap$, return map $\retmap$, and entry
map $\entmap$): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state $\mpstatename$ and any adjacent pair
of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
trace from $\mpstatename$ whose current instruction
belongs to different functions according to the function map,
$\mach_1$ is either a valid call or a valid return:

@@

For example, consider changing line 5 of the running example to a $\jalr$,
$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
to jump into the middle of {\tt h}. 
This violates control separation because the instruction that causes a
function switch is not a valid call (or a return) according to the
call map.

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state $\mpstatename$
and any n-transition \((\mach_c,\pol_c)\) and
\((\mach_e, \pol_e)\) in the induced trace from
$\mpstatename$, $\mach_e$ is a member of the entry map:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

The changed program described just above also violates entry integrity,
because the target (instruction 42) of the new $\jalr$ is not an established entry point
according to the entry map.

@@

\newcommand*{\Unmatched}{\mathit{um}}

Return integrity captures the expected behavior of marked returns with
respect to their call points. Intuitively, it says that the marked return
matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is a call state then
the matching return will be the first \emph{unmatched} return in $\MPT$. So
we begin by defining an auxiliary ``unmatched'' relation between two MP-traces.

\definition
We write $\Unmatched~\MPT~\MPT'$ if $\MPT'$ is
obtained from $\MPT$ by dropping elements until we reach the first unmatched return
(according to the return map).  Formally, $\Unmatched$ is defined coinductively as follows:
\begin{itemize}
\item A trace that starts with a return is related to itself:

\begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
\end{minipage}

@@

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))
              \hspace{.5cm}\Unmatched ~ \MPT ~ \MPT'\)}
           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}

\item Finally, if a trace starts with a call, then we must discard the first
  unmatched return from its tail to get to the real one:

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
\apt{Could cut formal definition for space.}

\definition%{Return Integrity}
A system enjoys {\em return integrity}
% with respect to a call map \(\callmap\) and return map \(\retmap\)
if, for any initial state pair 
$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
is the first unmatched return in $\MPT$, then $\mach_r$ corresponds
to a return state for $\mach_c$:

\[
\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
\]
\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
  And what about the no-unmatched-return-at-top-level clause?}

For example, if we modified the running example by adding
an instruction $ \rra \rightarrow \rra - 19$
between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
return to {\tt main} instead of {\tt g}. While the instruction that
causes the return would still be a valid return (satisfying control
separation), the state after the return would not correspond to an
unmatched return for {\tt g}'s call to {\tt h}, violating return
integrity.

@@

The above properties are somewhat orthogonal, but together they
capture a standard notion of control flow for a stack: functions keep
their code separate outside of calls and returns, calls respect
explicit entry points, and returns respect implicit entry points
corresponding to their call.  A system that enjoys all three
properties (with respect to a function map, call map, return map,
and entry map) is said to enjoy
{\em well-bracketed control flow} (with respect to those maps).

@@

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
  and stack confidentiality test $\testConf{\mach}{\mach'}$
  with respect to the original machine state. We write
  this $\testVS{(\mach, \pol)}{\vs}$.

@@

\definition
%
  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
  satisfies the eager testing property at depth 1 with respect to the initial
  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
%
  \rb{I don't think the initial variant has been defined, but it's
    straightforward.}

@@

      A system enjoys {\em observable stack integrity} with respect to
      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with respect to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with respect to a call map
      and a function map if it enjoys both observable stack integrity and
      observable stack confidentiality with respect to them.
